---
title: "Modeling"
author: "Rachel"
date: "7/16/2021"
output: html_document
---

Load libraries:
```{r}
library(tidyverse)
library(dplyr)
library(boot)
library(factoextra)
library(ggfortify)
library(corrplot)
library(psych)
library(ggplot2)
library(ggfortify)
library(cluster)
library(ggforce)
library(extrafont)
library(ggpubr)
library(robustbase)
library(MASS)
library(car)
library(pwr)
library(kinship2)
library(coxme)
library(gaston)
library(ggeffects)
```

Load data:
```{r}
Full_Het_df<-read.csv("Full_heterozygosity_df.csv")
MHC_ddRAD_Similarity_tidy<- read.csv("MHC_ddRAD_Similarity_tidy.csv")
```

create kinship matrix
```{r}
bed <- read.bed.matrix("refmap_pruned")
kin <- gaston::GRM(bed, autosome.only = FALSE)
rownames(kin) <- c(1:22)
colnames(kin) <- c(1:22)
```

power analysis
numerator degrees of freedom= # of predictors
denominator degrees of freedom= # of observations - (# of predictors +1)
```{r}
pwr.f2.test(u = 1, v = 20, sig.level = 0.05, power =0.8 )
pwr.f2.test(u = 1, v = 20, sig.level = 0.05, power =0.5 )
pwr.f2.test(u = 1, v = 20, sig.level = 0.05, power= 0.1 )
```

**Variable KEY includes description of columns in dataset** 

#Aim 3: Relationship between Class I and II diversity
```{r}
mod10 <- lmekin(MHCII ~ MHCAB + (1|ID), data=Full_Het_df, varlist=kin); mod10
mod11 <- lmekin(ClassII_distinct_supertypes ~ AB_distinct_supertypes + (1|ID), data=Full_Het_df, varlist=kin); mod11
```


# Aim 4: genome-wide vs MHC heterozygosity models
```{r}
mod1 <- lmekin(MHCAB ~ stMLH + (1|ID), data=Full_Het_df, varlist=kin); mod1
mod2 <- lmekin(MHCII ~ stMLH + (1|ID), data=Full_Het_df, varlist=kin); mod2
mod3 <- lmekin(ABMeanDiff ~ stMLH + (1|ID), data=Full_Het_df, varlist=kin); mod3
mod4 <- lmekin(DQAMeanDiff ~ stMLH + (1|ID), data=Full_Het_df, varlist=kin); mod4
mod5 <- lmekin(DRBMeanDiff ~ stMLH + (1|ID), data=Full_Het_df, varlist=kin); mod5
mod6 <- lmekin(AB_distinct_supertypes ~ stMLH + (1|ID), data=Full_Het_df, varlist=kin); mod6
mod7 <- lmekin(ClassII_distinct_supertypes ~ stMLH + (1|ID), data=Full_Het_df, varlist=kin); mod7
```

```{r}
p.adjust(c(0.36, 0.35, 0.97, 0.19, 0.19, 0.49, 0.68), method="BH")
```

# Using absolute heterozygosity instead of stMLH
```{r}
mod1abs <- lmekin(MHCAB ~ Refmap + (1|ID), data=Full_Het_df, varlist=kin); mod1abs
mod2abs <- lmekin(MHCII ~ Refmap + (1|ID), data=Full_Het_df, varlist=kin); mod2abs
mod3abs <- lmekin(ABMeanDiff ~ Refmap + (1|ID), data=Full_Het_df, varlist=kin); mod3abs
mod4abs <- lmekin(DQAMeanDiff ~ Refmap + (1|ID), data=Full_Het_df, varlist=kin); mod4abs
mod5abs <- lmekin(DRBMeanDiff ~ Refmap + (1|ID), data=Full_Het_df, varlist=kin); mod5abs
mod6abs <- lmekin(AB_distinct_supertypes ~ Refmap + (1|ID), data=Full_Het_df, varlist=kin); mod6abs
mod7abs <- lmekin(ClassII_distinct_supertypes ~ Refmap + (1|ID), data=Full_Het_df, varlist=kin); mod7abs
```

```{r}
p.adjust(c(0.26, 0.44, 0.95, 0.21, 0.13, 0.56, 0.92), method="BH")
```





# Aim 5: ddRAD and MHC complementarity robust models
```{r}
ClassI_fullmod <- glmrob(MHCI_common ~ KINSHIP + MHCI_unique, data= MHC_ddRAD_Similarity_tidy, family=poisson);
ClassI_redmod <- glmrob(MHCI_common ~ MHCI_unique, data= MHC_ddRAD_Similarity_tidy, family=poisson);

kinshipmod1 <- anova(ClassI_fullmod, ClassI_redmod)
summary(ClassI_fullmod)
```

```{r}
ClassII_fullmod <- glmrob(MHCII_common ~ KINSHIP + MHCII_unique, data= MHC_ddRAD_Similarity_tidy, family=poisson);
ClassII_redmod <- glmrob(MHCII_common ~ MHCII_unique, data= MHC_ddRAD_Similarity_tidy, family=poisson);

kinshipmod2 <- anova(ClassII_fullmod, ClassII_redmod)
summary(ClassII_fullmod)
```


```{r}
SupertypeI_fullmod <- glmrob(SupertypeAB_common ~ KINSHIP + SupertypesAB_unique, data= MHC_ddRAD_Similarity_tidy, family= poisson)

SupertypeI_redmod <- glmrob(SupertypeAB_common ~ SupertypesAB_unique, data= MHC_ddRAD_Similarity_tidy, family= poisson)

kinshipmod3 <- anova(SupertypeI_fullmod, SupertypeI_redmod)
summary(SupertypeI_fullmod)
```


```{r}
SupertypeII_fullmod <- glmrob(SupertypeII_common ~ KINSHIP + SupertypesII_unique, data= MHC_ddRAD_Similarity_tidy, family= poisson)

SupertypeII_redmod <- glmrob(SupertypeII_common ~ SupertypesII_unique, data= MHC_ddRAD_Similarity_tidy, family= poisson)

kinshipmod4 <- anova(SupertypeII_fullmod, SupertypeII_redmod)
summary(SupertypeII_fullmod)
```

```{r}
p.adjust(c(kinshipmod1[2,4], kinshipmod2[2,4], kinshipmod3[2,4], kinshipmod4[2,4]), method="BH")
```
