---
title: "Supertypes"
author: "Rachel"
date: "7/16/2021"
output: html_document
---

Load libraries
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(dynamicTreeCut)
library(ggplot2)
library(ggtree)
library(rcartocolor)
```

Load data
```{r}
ABProteins<- read.csv("ABProteins.csv")
DQAProteins<- read.csv("DQAProteins.csv")
DRBProteins<- read.csv("DRBProteins.csv")
zscores <- read.csv("zscores.csv")
```

Make new column "together" in z-score data frame that contains all z-scores for an AA separated by a comma
```{r}
zscores$together <- paste(zscores$z1, zscores$z2, zscores$z3, zscores$z4, zscores$z5, sep = ",")
```


## AB supertypes

Unique AB alleles
```{r}
ABUniqueNamed <- ABProteins %>%
  select(-Number, -Copies, -ID, -rownames) %>%
  distinct()
```

Make new column with AA present at each positively selected site for each allele
```{r}
i <- 1
PSS <- c(16, 25, 27, 38, 41, 42, 44, 45, 48, 49, 51, 52, 54, 55, 57, 58)
ABUniqueNamed$PSSites <- "X"

for(i in 1:nrow(ABUniqueNamed)){
  for (j in PSS){
    AA <- str_sub(ABUniqueNamed[i,4], j, j)
    ABUniqueNamed[i,7] <- paste(ABUniqueNamed[i,7], AA)
  }
  i <- i+1
}
```

Remove leading "X" from that column
```{r}
AB_PSS_spaces <- ABUniqueNamed %>%
  mutate(PSSites= str_trim(PSSites)) %>%
  mutate(PSSites = str_sub(PSSites, 2, nchar(PSSites))) %>%
  mutate(Name= str_replace_all(Name, "[*:]", "_")) %>%
  select(Name, PSSites)
```

Separate AAs into their own columns
```{r}
AB_sep <- AB_PSS_spaces %>%
  separate(PSSites, c("x0","x1","x2","x3","x4","x5","x6","x7","x8","x9","x10","x11","x12","x13","x14","x15","x16")) %>%
  select(-"x0")

#View(AB_sep)
```

Replace each AA with its 5 z-scores
```{r}
colnum = c("x1","x2","x3","x4","x5","x6","x7","x8","x9","x10","x11","x12","x13","x14","x15","x16")

for (i in colnum) {
  AB_sep[AB_sep[[i]] == "A", i] <- zscores[1, 9]
  AB_sep[AB_sep[[i]] == "R", i] <- zscores[2, 9]
  AB_sep[AB_sep[[i]] == "N", i] <- zscores[3, 9]
  AB_sep[AB_sep[[i]] == "D", i] <- zscores[4, 9]
  AB_sep[AB_sep[[i]] == "C", i] <- zscores[5, 9]
  AB_sep[AB_sep[[i]] == "Q", i] <- zscores[6, 9]
  AB_sep[AB_sep[[i]] == "E", i] <- zscores[7, 9]
  AB_sep[AB_sep[[i]] == "G", i] <- zscores[8, 9]
  AB_sep[AB_sep[[i]] == "H", i] <- zscores[9, 9]
  AB_sep[AB_sep[[i]] == "I", i] <- zscores[10, 9]
  AB_sep[AB_sep[[i]] == "L", i] <- zscores[11, 9]
  AB_sep[AB_sep[[i]] == "K", i] <- zscores[12, 9]
  AB_sep[AB_sep[[i]] == "M", i] <- zscores[13, 9]
  AB_sep[AB_sep[[i]] == "F", i] <- zscores[14, 9]
  AB_sep[AB_sep[[i]] == "P", i] <- zscores[15, 9]
  AB_sep[AB_sep[[i]] == "S", i] <- zscores[16, 9]
  AB_sep[AB_sep[[i]] == "T", i] <- zscores[17, 9]
  AB_sep[AB_sep[[i]] == "W", i] <- zscores[18, 9]
  AB_sep[AB_sep[[i]] == "Y", i] <- zscores[19, 9]
  AB_sep[AB_sep[[i]] == "V", i] <- zscores[20, 9]
}

#View(AB_sep)
```

Separate each z-score into its own column
```{r}
for (i in colnum) {
  AB_sep <- AB_sep %>%
  separate(i, c(paste(i,"z1",sep=""), paste(i,"z2",sep=""), paste(i,"z3",sep=""), paste(i,"z4",sep=""), paste(i,"z5",sep="")), sep=",")
}

AB_genesis <- AB_sep %>%
  select(-Name) %>%
  mutate_if(is.character,as.numeric)

AB_genesis <- cbind(AB_sep$Name, AB_genesis) %>%
  rename(Name = "AB_sep$Name")
```

Allele name is row, z score columns, 5 for each AA
```{r}
#make first column into rownames
AB_hclust <- AB_genesis[,-1]
rownames(AB_hclust) <- AB_genesis[,1]

#create distance matrix and perform hierarchical clustering
ab_dM <- as.matrix(dist(AB_hclust, method = "euclidean"))
AB_dist <- as.dist(ab_dM)
ab_hc <- hclust(AB_dist, method="average")

#dyanamic cut
ab_cut <- cutreeDynamic(ab_hc, method = "hybrid", distM= ab_dM, minClusterSize = 2, deepSplit=2)

ab_supertypes <- as.data.frame(ab_cut)

ab_supertypes <- cbind(ab_supertypes, AB_genesis$Name) %>%
  rename(Name = "AB_genesis$Name") %>%
  rename(Supertype= "ab_cut") %>%
  select(Name, Supertype)
#View(ab_supertypes)
#dynamic cut into 12 supertypes
```

```{r}
saveRDS(ab_hc, file="ab_hc.Rdata")
write.csv(ab_supertypes, "ab_supertypes.csv", row.names = FALSE)
```

# DQA supertypes

Unique DQA alleles
```{r}
DQAUniqueNamed <- DQAProteins %>%
  select(-Number, -Copies, -ID, -rownames) %>%
  distinct()
```

Make new column with AA present at each positively selected site for each allele
```{r}
i <- 1
PSS <- c(10,58,65,68)
DQAUniqueNamed$PSSites <- "X"

for(i in 1:nrow(DQAUniqueNamed)){
  for (j in PSS){
    AA <- str_sub(DQAUniqueNamed[i,4], j, j)
    DQAUniqueNamed[i,7] <- paste(DQAUniqueNamed[i,7], AA)
  }
  i <- i+1
}
```

Remove leading "X" from that column
```{r}
DQA_PSS_spaces <- DQAUniqueNamed %>%
  mutate(PSSites= str_trim(PSSites)) %>%
  mutate(PSSites = str_sub(PSSites, 2, nchar(PSSites))) %>%
  mutate(Name= str_replace_all(Name, "[*:]", "_")) %>%
  select(Name, PSSites)
```

Separate AAs into their own columns
```{r}
DQA_sep <- DQA_PSS_spaces %>%
  separate(PSSites, c("x0","x1","x2","x3","x4")) %>%
  select(-"x0")
```

Replace each AA with its 5 z-scores
```{r}
colnum = c("x1","x2","x3","x4")

for (i in colnum) {
  DQA_sep[DQA_sep[[i]] == "A", i] <- zscores[1, 9]
  DQA_sep[DQA_sep[[i]] == "R", i] <- zscores[2, 9]
  DQA_sep[DQA_sep[[i]] == "N", i] <- zscores[3, 9]
  DQA_sep[DQA_sep[[i]] == "D", i] <- zscores[4, 9]
  DQA_sep[DQA_sep[[i]] == "C", i] <- zscores[5, 9]
  DQA_sep[DQA_sep[[i]] == "Q", i] <- zscores[6, 9]
  DQA_sep[DQA_sep[[i]] == "E", i] <- zscores[7, 9]
  DQA_sep[DQA_sep[[i]] == "G", i] <- zscores[8, 9]
  DQA_sep[DQA_sep[[i]] == "H", i] <- zscores[9, 9]
  DQA_sep[DQA_sep[[i]] == "I", i] <- zscores[10, 9]
  DQA_sep[DQA_sep[[i]] == "L", i] <- zscores[11, 9]
  DQA_sep[DQA_sep[[i]] == "K", i] <- zscores[12, 9]
  DQA_sep[DQA_sep[[i]] == "M", i] <- zscores[13, 9]
  DQA_sep[DQA_sep[[i]] == "F", i] <- zscores[14, 9]
  DQA_sep[DQA_sep[[i]] == "P", i] <- zscores[15, 9]
  DQA_sep[DQA_sep[[i]] == "S", i] <- zscores[16, 9]
  DQA_sep[DQA_sep[[i]] == "T", i] <- zscores[17, 9]
  DQA_sep[DQA_sep[[i]] == "W", i] <- zscores[18, 9]
  DQA_sep[DQA_sep[[i]] == "Y", i] <- zscores[19, 9]
  DQA_sep[DQA_sep[[i]] == "V", i] <- zscores[20, 9]
}

#View(DQA_sep)
```

Separate each z-score into its own column
```{r}
for (i in colnum) {
  DQA_sep <- DQA_sep %>%
  separate(i, c(paste(i,"z1",sep=""), paste(i,"z2",sep=""), paste(i,"z3",sep=""), paste(i,"z4",sep=""), paste(i,"z5",sep="")), sep=",")
}

#View(DQA_sep)

DQA_genesis <- DQA_sep %>%
  select(-Name) %>%
  mutate_if(is.character,as.numeric)

DQA_genesis <- cbind(DQA_sep$Name, DQA_genesis) %>%
  rename(Name = "DQA_sep$Name")
```

Allele name is row, z score columns, 5 for each AA
```{r}
#make first column into rowname 
DQA_hclust <- DQA_genesis[,-1]
rownames(DQA_hclust) <- DQA_genesis[,1]

#create distance matrix and perform hierarchical clustering
dqa_dM <- as.matrix(dist(DQA_hclust, method = "euclidean"))
DQA_dist <- as.dist(dqa_dM)
dqa_hc <- hclust(DQA_dist, method="average")

#dynamic cut tree into clusters
dqa_cut <- cutreeDynamic(dqa_hc, method = "hybrid", distM= dqa_dM, minClusterSize = 1, deepSplit=2)

dqa_supertypes <- as.data.frame(dqa_cut)

dqa_supertypes <- cbind(dqa_supertypes, DQA_genesis$Name) %>%
  rename(Name = "DQA_genesis$Name") %>%
  rename(Supertype= "dqa_cut") %>%
  select(Name, Supertype)
#View(dqa_supertypes)
#dynamic cut into 3 supertypes
```

```{r}
saveRDS(dqa_hc, "dqa_hc.Rdata")
write.csv(dqa_supertypes, "dqa_supertypes.csv", row.names = FALSE)
```


### DRB supertypes

Unique DRB alleles
```{r}
DRBUniqueNamed <- DRBProteins %>%
  select(-Number, -Copies, -ID, -rownames) %>%
  distinct()
```

Make new column with AA present at each positively selected site for each allele
```{r}
i <- 1
PSS <- c(4,6,19,21,23,25,30,31,50,53,57,60,63,64,67,71,77,79)
DRBUniqueNamed$PSSites <- "X"

for(i in 1:nrow(DRBUniqueNamed)){
  for (j in PSS){
    AA <- str_sub(DRBUniqueNamed[i,4], j, j)
    DRBUniqueNamed[i,7] <- paste(DRBUniqueNamed[i,7], AA)
  }
  i <- i+1
}
```

Remove the leading X character
```{r}
DRB_PSS_spaces <- DRBUniqueNamed %>%
  mutate(PSSites= str_trim(PSSites)) %>%
  mutate(PSSites = str_sub(PSSites, 2, nchar(PSSites))) %>%
  mutate(Name= str_replace_all(Name, "[*:]", "_")) %>%
  select(Name, PSSites)
```

Separate each AA into its own column
```{r}
DRB_sep <- DRB_PSS_spaces %>%
  separate(PSSites, c("x0","x1","x2","x3","x4","x5","x6","x7","x8","x9","x10","x11","x12","x13","x14","x15","x16", "x17", "x18")) %>%
  select(-"x0")

#View(DRB_sep)
```

Replace each AA with its 5 z-scores
```{r}
colnum = c("x1","x2","x3","x4","x5","x6","x7","x8","x9","x10","x11","x12","x13","x14","x15","x16", "x17", "x18")

for (i in colnum) {
  DRB_sep[DRB_sep[[i]] == "A", i] <- zscores[1, 9]
  DRB_sep[DRB_sep[[i]] == "R", i] <- zscores[2, 9]
  DRB_sep[DRB_sep[[i]] == "N", i] <- zscores[3, 9]
  DRB_sep[DRB_sep[[i]] == "D", i] <- zscores[4, 9]
  DRB_sep[DRB_sep[[i]] == "C", i] <- zscores[5, 9]
  DRB_sep[DRB_sep[[i]] == "Q", i] <- zscores[6, 9]
  DRB_sep[DRB_sep[[i]] == "E", i] <- zscores[7, 9]
  DRB_sep[DRB_sep[[i]] == "G", i] <- zscores[8, 9]
  DRB_sep[DRB_sep[[i]] == "H", i] <- zscores[9, 9]
  DRB_sep[DRB_sep[[i]] == "I", i] <- zscores[10, 9]
  DRB_sep[DRB_sep[[i]] == "L", i] <- zscores[11, 9]
  DRB_sep[DRB_sep[[i]] == "K", i] <- zscores[12, 9]
  DRB_sep[DRB_sep[[i]] == "M", i] <- zscores[13, 9]
  DRB_sep[DRB_sep[[i]] == "F", i] <- zscores[14, 9]
  DRB_sep[DRB_sep[[i]] == "P", i] <- zscores[15, 9]
  DRB_sep[DRB_sep[[i]] == "S", i] <- zscores[16, 9]
  DRB_sep[DRB_sep[[i]] == "T", i] <- zscores[17, 9]
  DRB_sep[DRB_sep[[i]] == "W", i] <- zscores[18, 9]
  DRB_sep[DRB_sep[[i]] == "Y", i] <- zscores[19, 9]
  DRB_sep[DRB_sep[[i]] == "V", i] <- zscores[20, 9]
}

#View(DRB_sep)
```

Separate each z-score into its own column
```{r}
for (i in colnum) {
  DRB_sep <- DRB_sep %>%
  separate(i, c(paste(i,"z1",sep=""), paste(i,"z2",sep=""), paste(i,"z3",sep=""), paste(i,"z4",sep=""), paste(i,"z5",sep="")), sep=",")
}

DRB_genesis <- DRB_sep %>%
  select(-Name) %>%
  mutate_if(is.character,as.numeric)

DRB_genesis <- cbind(DRB_sep$Name, DRB_genesis) %>%
  rename(Name = "DRB_sep$Name")
```

Allele name is row, z-score columns, 5 for each AA
```{r}
#make first column the rownames
DRB_hclust <- DRB_genesis[,-1]
rownames(DRB_hclust) <- DRB_genesis[,1]

#create distance matrix and perform hierarchical clustering
drb_dM <- as.matrix(dist(DRB_hclust, method = "euclidean"))
DRB_dist <- as.dist(drb_dM)
drb_hc <- hclust(DRB_dist, method="average")

#dynamic cut tree into clusters
drb_cut <- cutreeDynamic(drb_hc, method = "hybrid", distM= drb_dM, minClusterSize = 2, deepSplit=2)

drb_supertypes <- as.data.frame(drb_cut)

drb_supertypes <- cbind(drb_supertypes, DRB_genesis$Name) %>%
  rename(Name = "DRB_genesis$Name") %>%
  rename(Supertype= "drb_cut") %>%
  select(Name, Supertype)
#View(drb_supertypes)
#dynamic cut into 6 supertypes
```

```{r}
saveRDS(drb_hc, "drb_hc.Rdata")
write.csv(drb_supertypes, "drb_supertypes.csv", row.names = FALSE)
```
