---
title: "Figures"
author: "Rachel"
date: "7/9/2021"
output: html_document
---

Load libraries

```{r}
library(tidyverse)
library(dplyr)
library(extrafont)
library(gdata)
library(ggplot2)
library(stringr)
library(ggtree)
library(rcartocolor)
library(ggpubr)
```

Load data:
```{r}
Full_Het_df<- read.csv("Full_heterozygosity_df.csv")
mapping_rate<-read.csv("mapping_rate.csv", header=TRUE)
ReadCutoffs <- read.csv("ReadCutoffs.csv")
drb_supertypes <- read.csv("drb_supertypes.csv")
dqa_supertypes <- read.csv("dqa_supertypes.csv")
ab_supertypes <- read.csv("ab_supertypes.csv")
ab_hc <- readRDS("ab_hc.Rdata")
dqa_hc <- readRDS("dqa_hc.Rdata")
drb_hc <- readRDS("drb_hc.Rdata")
```


Aim 1: Characterize global heterozygosity and kinship 

Figure S2: ddRADseq mapping rate
```{r}
mapping_rate$ID <- as.character(mapping_rate$ID)

Mapping<- ggplot(mapping_rate, aes(x=reorder(ID, -Mapped), y=Mapped)) + 
  geom_col(fill= "grey41") +
  scale_y_continuous(name="Number of Mapped Reads", labels = scales::unit_format(unit = "M", scale = 1e-6), expand = expansion(mult = c(0, .02))) +
  xlab("Animal ID") +
    scale_x_discrete(labels= c("1", "2", "21", "16", "3", "6", "15", "11", "14", "20", "7", "9", "8", "13", "19", "17", "10", "5", "18", "4", "12", "22")) +
  theme_light() +
  theme(text=element_text(family="Times New Roman")) +
  theme(axis.text.x = element_text(angle = 90, size=6, hjust = 1, vjust = 0.5)); Mapping

ggsave(Mapping,file='Mapping.pdf',width=8,height=4, device=cairo_pdf)
```

Figure S3: Heatmap of relatedness
```{r}
IBD <- read.csv('IBDinput.csv', h=T)
id_map <- read.csv("id_map.csv", h=T)

id_map <- id_map %>%
  mutate(Name1 = ID) %>%
  mutate(Name2= ID)

IBD_tidy <- left_join(IBD, id_map, by= "Name1")

IBD_tidy <- IBD_tidy %>%
  dplyr::select(Name1, Name2.x, IBD, ID1, ID2, ID_number) %>%
  mutate(Name2= Name2.x) %>%
  mutate(ID_1 = ID_number)

IBD_tidy <- left_join(IBD_tidy, id_map, by= "Name2")

IBD_tidy <- IBD_tidy %>%
  mutate(ID_2= ID_number.y) %>%
  mutate(Name1= Name1.x) %>%
  dplyr::select(Name1, Name2, IBD, ID_1, ID_2) %>%
  mutate(ID1= ID_1) %>%
  mutate(ID2= ID_2) %>%
  dplyr::select(Name1, Name2, IBD, ID1, ID2)

IBD_tidy$ID1 <- as.character(IBD_tidy$ID1)
IBD_tidy$ID2 <- as.character(IBD_tidy$ID2)

IBD1 <- subset(IBD_tidy, IBD==1)

KIN_heat <- ggplot(data=IBD_tidy, aes(x=ID1, y=ID2, fill=IBD)) + 
  geom_tile() + 
  geom_tile(data=IBD1, fill="black") +
  scale_fill_gradient2(low = "grey100", mid= "grey73", high = "grey44", midpoint = 0.25, limit = c(0,0.5), name= "Relatedness") +
  scale_color_manual(IBD==1, values = "white")+
  geom_text(aes(ID1, ID2, label = round(IBD,2)), color = "black", size = 2.2, family= "Times New Roman") +
  scale_x_discrete(limits=c("1", "2", "3","4", "5", "6","7", "8", "9","10", "11", "12","13", "14", "15","16", "17", "18","19", "20", "21", "22")) +
  scale_y_discrete(limits=c("1", "2", "3","4", "5", "6","7", "8", "9","10", "11", "12","13", "14", "15","16", "17", "18","19", "20", "21", "22"))+
  xlab("Baboon ID") +
  ylab("Baboon ID") +
  theme(text=element_text(family="Times New Roman"), legend.text = element_text(size = 7)); KIN_heat

ggsave("KIN_heat.pdf", KIN_heat, width= 200, height= 150, units= "mm", device = cairo_pdf)
```


Aim 2: Characterize MHC heterozygosity and complementarity

Figure S1: Mean intra-individual drops in sequence copy number 
```{r}
ReadCutoffs_wide <- ReadCutoffs %>%
  pivot_wider(names_from= Condition, values_from= Reads) %>%
  mutate(Difference = (above-below)) %>%
  mutate(Multiple = (above/below))

Difference_plot <- ggplot(ReadCutoffs_wide, aes(Loci, Difference)) +
  geom_boxplot(fill= "grey") +
  theme_classic() +
  theme(text=element_text(family="Times New Roman")) +
  
  scale_y_continuous(trans='log10', labels= scales::comma) +
  ylab("Difference in copy number between retained and filtered sequences"); Difference_plot

ggsave(Difference_plot,file='ReadCutoffPlot.pdf',width=5,height=5, device = cairo_pdf)
```
 

Figure 1: Frequency distribution of MHC alleles and supertypes
```{r}
MHCIhist <- ggplot(aes(x= MHCAB), data=Full_Het_df) +
  geom_bar(color="black", fill="white") +
  theme_light() +
  theme(text=element_text(family="Times New Roman")) +
  scale_x_continuous(breaks= c(1,2,3,4,5,6,7,8,9,10,11,12,13,14), limits= c(7,14)) +
  scale_y_continuous(breaks= c(0,1,2,3,4,5,6,7,8,9,10)) +
  xlab("Number of Class I alleles") +
  ylab("Number of baboons")

MHCIIhist <- ggplot(aes(x= MHCII), data=Full_Het_df) +
  geom_bar(color="black", fill="white") +
  theme_light() +
  theme(text=element_text(family="Times New Roman")) +
  scale_x_continuous(breaks= c(1,2,3,4,5,6,7,8,9,10)) +
  scale_y_continuous(breaks= c(0,1,2,3,4,5,6,7,8,9,10)) +
  xlab("Number of Class II alleles") +
  ylab("")

MHCIsuperhist <- ggplot(aes(x= AB_distinct_supertypes), data=Full_Het_df) +
  geom_bar(color="black", fill="white") +
  theme_light() +
  theme(text=element_text(family="Times New Roman")) +
  scale_x_continuous(breaks= c(4,5,6,7,8,9,10,11), limits= c(4,11)) +
  scale_y_continuous(breaks= c(0,1,2,3,4,5,6,7), limits= c(0,6)) +
  xlab("Number of Class I supertypes") +
  ylab("Number of baboons")

MHCIIsuperhist <- ggplot(aes(x= ClassII_distinct_supertypes), data=Full_Het_df) +
  geom_bar(color="black", fill="white") +
  theme_light() +
  theme(text=element_text(family="Times New Roman")) +
  scale_x_continuous(breaks= c(1,2,3,4,5,6,7,8), limits= c(1,8)) +
  scale_y_continuous(breaks= c(0,1,2,3,4,5,6,7,8,9,10)) +
  xlab("Number of Class II supertypes") +
  ylab("")

MHCdiversityhist <- ggarrange(MHCIhist, MHCIIhist, MHCIsuperhist, MHCIIsuperhist, ncol=2, nrow= 2, labels= "auto", font.label=list(size= 14, family="Times New Roman")); MHCdiversityhist

ggsave("Fig1.MHCHist.pdf", MHCdiversityhist, height= 150, width= 150, units= "mm", device = cairo_pdf)
```

Figure S5: Twelve MHC A & B supertypes
```{r}
tree <- ggtree(ab_hc) +
  geom_tiplab(offset= 0.2, cex=2.5, family="Times New Roman") +
  xlim(0, 14)

rotatetree <- tree %>%
  flip(62,63) %>%
  flip(68,69) %>%
  flip(39,72) %>%
  flip(32,54)

tiptree <- rotatetree %<+% ab_supertypes + 
       geom_tippoint(aes(color=factor(Supertype), shape=factor(Supertype)), size=4) +
  scale_color_carto_d(palette = "Safe") +
  scale_shape_manual(values=rep(15:19, 3)) +
  labs(color= "Supertype", shape="Supertype") +
  theme(text=element_text(family="Times New Roman"))
tiptree

ggsave("AB_supertype_tree.pdf", tiptree, height=200, width= 250, units= "mm", device= cairo_pdf)
```

Figure S6: Three MHC DQA supertypes
```{r}
tree <- ggtree(dqa_hc) +
  geom_tiplab(offset= 0.2, cex=3, family= "Times New Roman") +
  xlim(0, 10)

rotatetree <- tree %>%
  flip(15, 16)

dqa_tiptree <- rotatetree %<+% dqa_supertypes + 
       geom_tippoint(aes(color=factor(Supertype), shape=factor(Supertype)), size=4) +
  scale_color_carto_d(palette = "Safe") +
  scale_shape_manual(values=c(15:17)) +
  labs(color= "Supertype", shape="Supertype") +
  theme(text=element_text(family="Times New Roman")); dqa_tiptree

ggsave("DQA_supertype_tree.pdf", dqa_tiptree, height= 100, width= 175, units= "mm", device= cairo_pdf)
```

Figure S9: Six MHC DRB supertypes
```{r}
tree <- ggtree(drb_hc) +
  geom_tiplab(offset= 0.2, cex=3, family="Times New Roman") +
  xlim(0, 16)

rotatetree <- tree %>%
  flip(44, 45) %>%
  flip(10, 33)

drb_tiptree <- rotatetree %<+% drb_supertypes + 
       geom_tippoint(aes(color=factor(Supertype), shape=factor(Supertype)), size=4) +
  scale_color_carto_d(palette = "Safe") +
  scale_shape_manual(values=rep(15:19, 2)) +
  labs(color= "Supertype", shape="Supertype") +
  theme(text=element_text(family="Times New Roman")); drb_tiptree

ggsave("DRB_supertype_tree.pdf", drb_tiptree, height= 150, width= 250, units= "mm", device=cairo_pdf)
```


Aim 3: Trade-offs in MHC diversity components

Figure 2: Negative relationship between class I and II diversity
```{r}
Supertypes <- ggplot(Full_Het_df,aes(y=ClassII_distinct_supertypes,x=AB_distinct_supertypes)) + 
  geom_point(shape= 18, color="black") + 
  geom_smooth(method="lm", color="black", alpha= 0.2, size=0.3) +
  theme_classic() +
  theme(text=element_text(family="Times New Roman")) +
  ylab("Number Class II Supertypes") +
  xlab("Number Class I Supertypes") 

Alleles <- ggplot(Full_Het_df,aes(y=MHCII,x=MHCAB)) + 
  geom_point(shape= 18, color="black") + 
  geom_smooth(method="lm", color="black", alpha=0.2, size=0.3) +
  theme_classic() + 
  theme(text=element_text(family="Times New Roman")) +
  ylab("Number Class II Alleles") +
  xlab("Number Class I Alleles") 

Aim3 <- ggarrange(Alleles, Supertypes, ncol=2, nrow=1)
ggsave("Aim3.pdf", Aim3, width=7, height=4, device = cairo_pdf)
```

