---
title: "Modeling"
author: "Rachel"
date: "7/16/2021"
output: html_document
---

Load libraries:
```{r}
library(tidyverse)
library(dplyr)
library(boot)
library(factoextra)
library(ggfortify)
library(corrplot)
library(psych)
library(ggplot2)
library(ggfortify)
library(cluster)
library(ggforce)
library(extrafont)
library(ggpubr)
library(robustbase)
library(MASS)
library(car)
library(pwr)
library(kinship2)
library(coxme)
library(gaston)
library(ggeffects)
```

Load data:
```{r}
Full_Het_df<-read.csv("Full_heterozygosity_df.csv")
MHC_ddRAD_Similarity_tidy<- read.csv("MHC_ddRAD_Similarity_tidy.csv")
```

Move each variable into its own object
```{r}
MHCAllhet <- Full_Het_df$MHC_All
MHCIhet <- Full_Het_df$MHCAB
MHCIIhet <- Full_Het_df$MHCII
MHCDQAhet <- Full_Het_df$MHCDQA
MHCDRBhet <- Full_Het_df$MHCDRB
SNPhet <- Full_Het_df$stMLH
ABMeanDiff <- Full_Het_df$ABMeanDiff
DQAMeanDiff <- Full_Het_df$DQAMeanDiff
DRBMeanDiff <- Full_Het_df$DRBMeanDiff
ABMinDiff <- Full_Het_df$ABMinDiff
DQAMinDiff <- Full_Het_df$DQAMinDiff
DRBMinDiff <- Full_Het_df$DRBMinDiff
ABMaxDiff <- Full_Het_df$ABMaxDiff
DQAMaxDiff <- Full_Het_df$DQAMaxDiff
DRBMaxDiff <- Full_Het_df$DRBMaxDiff
AllSuper <- Full_Het_df$All_distinct_supertypes
ABSuper <- Full_Het_df$AB_distinct_supertypes
DQASuper <- Full_Het_df$DQA_distinct_supertypes
DRBSuper <- Full_Het_df$DRB_distinct_supertypes
ClassIISuper <- Full_Het_df$ClassII_distinct_supertypes
```


#Aim 3: Relationship between Class I and II diversity robust models
```{r}
mod10 <- rlm(ClassIISuper ~ ABSuper)
summary(mod10)
Anova(mod10, type= "III")


mod11 <- rlm(MHCIIhet ~ MHCIhet)
summary(mod11)
Anova(mod11, type= "III")
```

```{r}
mod10kin <- lmekin(ClassII_distinct_supertypes ~ AB_distinct_supertypes + (1|ID), data=Full_Het_df, varlist=kin); mod10kin

mod11kin <- lmekin(MHCII ~ MHCAB + (1|ID), data=Full_Het_df, varlist=kin); mod11kin
```


# Aim 4: ddRAD and MHC heterozygosity kinship models

```{r}
mod1 <- rlm(MHCIhet ~ SNPhet)
summary(mod1)
Anovamod1 <- Anova(mod1, type= "III"); Anovamod1

mod2 <- rlm(MHCIIhet ~ SNPhet)
summary(mod2)
Anovamod2 <- Anova(mod2, type= "III"); Anovamod2

mod3 <- rlm(ABMeanDiff ~ SNPhet)
summary(mod3)
Anovamod3 <- Anova(mod3, type= "III"); Anovamod3


mod4 <- rlm(DQAMeanDiff ~ SNPhet)
summary(mod4)
Anovamod4 <- Anova(mod4, type= "III"); Anovamod4

mod5 <- rlm(DRBMeanDiff~ SNPhet)
summary(mod5)
Anovamod5 <- Anova(mod5, type= "III"); Anovamod5

mod6 <- rlm(ABSuper ~ SNPhet)
summary(mod6)
Anovamod6 <- Anova(mod6, type= "III"); Anovamod6

mod7 <- rlm(ClassIISuper ~ SNPhet)
summary(mod7)
Anovamod7 <- Anova(mod7, type= "III"); Anovamod7


p.adjust(c(Anovamod1[2,3], Anovamod2[2,3], Anovamod3[2,3], Anovamod4[2,3], Anovamod5[2,3], Anovamod6[2,3], Anovamod7[2,3]), method= "BH")
```





kinship analyses
```{r}
bed <- read.bed.matrix("refmap_pruned")
kin <- gaston::GRM(bed, autosome.only = FALSE)
rownames(kin) <- c(1:22)
colnames(kin) <- c(1:22)
```

```{r}
mod1kin <- lmekin(MHCAB ~ stMLH + (1|ID), data=Full_Het_df, varlist=kin); mod1kin

mod2kin <- lmekin(MHCII ~ stMLH + (1|ID), data=Full_Het_df, varlist=kin); mod2kin

mod3kin <- lmekin(ABMeanDiff ~ stMLH + (1|ID), data=Full_Het_df, varlist=kin); mod3kin

mod4kin <- lmekin(DQAMeanDiff ~ stMLH + (1|ID), data=Full_Het_df, varlist=kin); mod4kin

mod5kin <- lmekin(DRBMeanDiff ~ stMLH + (1|ID), data=Full_Het_df, varlist=kin); mod5kin

mod6kin <- lmekin(AB_distinct_supertypes ~ stMLH + (1|ID), data=Full_Het_df, varlist=kin); mod6kin

mod7kin <- lmekin(ClassII_distinct_supertypes ~ stMLH + (1|ID), data=Full_Het_df, varlist=kin); mod7kin
```

power analysis
numerator degrees of freedom= # of predictors
demoninator degrees of freedom= # of observations - (# of predictors +1)
```{r}
pwr.f2.test(u = 1, v = 20, sig.level = 0.05, power =0.8 )

pwr.f2.test(u = 1, v = 20, sig.level = 0.05, power =0.5 )

pwr.f2.test(u = 1, v = 20, sig.level = 0.05, power= 0.1 )
```



# Aim 5: ddRAD and MHC complementarity robust models

```{r}
ClassI_fullmod <- glmrob(MHCI_common ~ KINSHIP + MHCI_unique, data= MHC_ddRAD_Similarity_tidy, family=poisson);
ClassI_redmod <- glmrob(MHCI_common ~ MHCI_unique, data= MHC_ddRAD_Similarity_tidy, family=poisson);

kinshipmod1 <- anova(ClassI_fullmod, ClassI_redmod)
summary(ClassI_fullmod)
```

```{r}
ClassII_fullmod <- glmrob(MHCII_common ~ KINSHIP + MHCII_unique, data= MHC_ddRAD_Similarity_tidy, family=poisson);
ClassII_redmod <- glmrob(MHCII_common ~ MHCII_unique, data= MHC_ddRAD_Similarity_tidy, family=poisson);

kinshipmod2 <- anova(ClassII_fullmod, ClassII_redmod)
summary(ClassII_fullmod)
```


```{r}
SupertypeI_fullmod <- glmrob(SupertypeAB_common ~ KINSHIP + SupertypesAB_unique, data= MHC_ddRAD_Similarity_tidy, family= poisson)

SupertypeI_redmod <- glmrob(SupertypeAB_common ~ SupertypesAB_unique, data= MHC_ddRAD_Similarity_tidy, family= poisson)

kinshipmod3 <- anova(SupertypeI_fullmod, SupertypeI_redmod)
summary(SupertypeI_fullmod)
```


```{r}
SupertypeII_fullmod <- glmrob(SupertypeII_common ~ KINSHIP + SupertypesII_unique, data= MHC_ddRAD_Similarity_tidy, family= poisson)

SupertypeII_redmod <- glmrob(SupertypeII_common ~ SupertypesII_unique, data= MHC_ddRAD_Similarity_tidy, family= poisson)

kinshipmod4 <- anova(SupertypeII_fullmod, SupertypeII_redmod)
summary(SupertypeII_fullmod)
```

```{r}
p.adjust(c(kinshipmod1[2,4], kinshipmod2[2,4], kinshipmod3[2,4], kinshipmod4[2,4]), method="fdr")
```



